shrimp_goby_boatnoise

### 2025-03-05 ms_boatnoise_shrimp_goby.Rmd
addressing Jake Martin's comments with addition of temporary code for explination

🚧 Interesting that a 16% of the observations are not being predicted very well by the model, do we think this is an issue 🚧

Using leave one out (loo) measure of fit, the model appears to preform ok, 16% Pareto k estimates falls outside the good range (0.7, 1]

```{r}
loo(goby_out_model)
```

@@

```{r}
pp_check(goby_out_model, ndraws = 50)
pp_check(goby_out_model_simple1, ndraws = 50)
pp_check(goby_out_model_simple2, ndraws = 50)
pp_check(goby_out_model_simple3, ndraws = 50)
pp_check(goby_out_model_simple4, ndraws = 50)
pp_check(goby_out_model, type='stat', stat='mean')
pp_check(goby_out_model_simple1, type='stat', stat='mean')
pp_check(goby_out_model_simple2, type='stat', stat='mean')
pp_check(goby_out_model_simple3, type='stat', stat='mean')
pp_check(goby_out_model_simple4, type='stat', stat='mean')
```

```{r}
loo(goby_out_model_simple1)
```

```{r}
loo(goby_out_model_simple2)
```

```{r}
loo(goby_out_model_simple3)
```

```{r}
loo(goby_out_model_simple4)
```

>>>>

@@

```{r}
pp_check(goby_out_model, ndraws = 50) + ggtitle("Main Model - Density Plot")
pp_check(goby_out_model_simple1, ndraws = 50) + ggtitle("Simple Model 1 - Density Plot")
pp_check(goby_out_model_simple2, ndraws = 50) + ggtitle("Simple Model 2 - Density Plot")
pp_check(goby_out_model_simple3, ndraws = 50) + ggtitle("Simple Model 3 - Density Plot")
pp_check(goby_out_model_simple4, ndraws = 50) + ggtitle("Simple Model 4 - Density Plot")
pp_check(goby_out_model, type = 'stat', stat = 'mean') + ggtitle("Main Model - Mean Check")
pp_check(goby_out_model_simple1, type = 'stat', stat = 'mean') + ggtitle("Simple Model 1 - Mean Check")
pp_check(goby_out_model_simple2, type = 'stat', stat = 'mean') + ggtitle("Simple Model 2 - Mean Check")
pp_check(goby_out_model_simple3, type = 'stat', stat = 'mean') + ggtitle("Simple Model 3 - Mean Check")
pp_check(goby_out_model_simple4, type = 'stat', stat = 'mean') + ggtitle("Simple Model 4 - Mean Check")
```

```{r}
loo(goby_out_model)
loo(goby_out_model_simple1)
loo(goby_out_model_simple2)
loo(goby_out_model_simple3)
loo(goby_out_model_simple4)
#I have nicely pasted outputs below in text
```

**Main model**
Pareto k diagnostic values:
                         Count Pct.    Min. ESS
(-Inf, 0.7]   (good)     310   84.0%   530     
   (0.7, 1]   (bad)       57   15.4%   <NA>    
   (1, Inf)   (very bad)   2    0.5%   <NA>

**Simple model1**
Pareto k diagnostic values:
                         Count Pct.    Min. ESS
(-Inf, 0.7]   (good)     341   92.4%   266     
   (0.7, 1]   (bad)       26    7.0%   <NA>    
   (1, Inf)   (very bad)   2    0.5%   <NA>

**Simple model2**
Pareto k diagnostic values:
                         Count Pct.    Min. ESS
(-Inf, 0.7]   (good)     341   92.4%   266     
   (0.7, 1]   (bad)       26    7.0%   <NA>    
   (1, Inf)   (very bad)   2    0.5%   <NA>

**Simple model3**
Pareto k diagnostic values:
                         Count Pct.    Min. ESS
(-Inf, 0.7]   (good)     349   94.6%   307     
   (0.7, 1]   (bad)       18    4.9%   <NA>    
   (1, Inf)   (very bad)   2    0.5%   <NA>

**Simple model4**
Pareto k diagnostic values:
                         Count Pct.    Min. ESS
(-Inf, 0.7]   (good)     367   99.5%   875     
   (0.7, 1]   (bad)        2    0.5%   <NA>    
   (1, Inf)   (very bad)   0    0.0%   <NA>

#results from rabbit hole

#################################################################################
Change **var_model** to suit the model you want to compare to the complete model
#################################################################################
```{r}
var_model = goby_out_model_simple2

var_model_contrasts <- var_model |>
  emmeans(~ stroke*stage, 
          epred = TRUE,
          re_formla = NULL,
          rg.limit =14000) |> 
  contrast(method = "revpairwise")

# Extract the relevant contrast information
var_model_contrasts <- as.data.frame(var_model_contrasts) |> 
  dplyr::filter(contrast %in% c("control Pre - 2 Pre", "control Pre - 4 Pre","4 Pre - 2 Pre", "control Pre - control During", "control Pre - control Post", "4 Pre - 4 During", "4 Pre - 4 Post", "2 Pre - 2 During", "2 Pre - 2 Post"))

var_model_contrasts %>% 
  gt()

var_model_estimates <- var_model |> 
  emmeans(~ stroke*stage,
          epred = TRUE,
          rg.limit = 14000,
          re_formula = NA)

var_model_estimates <- as.data.frame(var_model_estimates)

var_model_draws <- var_model |> 
  emmeans(~ stroke*stage,
          epred = TRUE,
          rg.limit = 14000,
          re_formula = NA) |> 
  gather_emmeans_draws()

var_model_draws <- var_model_draws |> 
  rename(emmean = .value)

var_model_estimates$stage <- factor(var_model_estimates$stage, levels = c("Pre", "During", "Post"))
var_model_estimates$stroke <- factor(var_model_estimates$stroke, levels = c("control", "4", "2"))

var_model_draws$stage <- factor(var_model_draws$stage, levels = c("Pre", "During", "Post"))
var_model_draws$stroke <- factor(var_model_draws$stroke, levels = c("control", "4", "2"))

var_fig_together <- ggplot(data = goby_out_raw_data, aes(x = stage, y = emmean, color = stroke, fill = stroke)) +
  geom_point(data = goby_out_means, alpha = 1, position = position_dodgenudge(width = 0.8, x = -0.09), size = 9, shape = "_") +
  geom_half_point(alpha = 0.3, position = position_dodgenudge(width = 0.8, x = -0.03), side = "l", range_scale = 0.5) +
  geom_half_violin(data = var_model_draws, aes(x = stage, y = emmean, color = stroke, fill = stroke), alpha = 0.6, position = position_dodge(width = 0.8), side = "r", size = 0) +
  geom_point(data = var_model_estimates, aes(x = stage, y = emmean, color = stroke), position = position_dodge(width = 0.8), size = 3) +
  geom_errorbar(data = var_model_estimates, aes(x = stage, y = emmean, ymin = lower.HPD, ymax = upper.HPD, color = stroke), position = position_dodge(width = 0.8), width = 0, size = 1.25) +
  scale_fill_manual(values = c("control" = "#76A34A", "4" = "#F29E00", "2" = "#E25A00"), name = "Noise exposure", guide = guide_legend(override.aes = list(size = 3))) +
  scale_color_manual(values = c("control" = "#2A330E", "4" = "#805213", "2" = "#542709"), name = "Noise exposure", guide = guide_legend(override.aes = list(size = 3))) +
  scale_x_discrete(labels = c("Pre", "During", "Post")) +
  labs(title = "The effects of boat noise on goby burrow use",
       y = "Proportion of time in out of the burrow",
       x = "Experimental phase relative to noise exposure") +
  theme_classic() +
  theme(
    axis.text.x = element_text(size = 10, margin = margin(t = 5)),  
    axis.text.y = element_text(size = 10, margin = margin(r = 5)),  
    axis.title.x = element_text(size = 13, margin = margin(t = 10)),
    axis.title.y = element_text(size = 13, margin = margin(r = 10)),
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 10),
    legend.position = "right",
    legend.justification = "top",
    plot.title = element_text(hjust = 0.5, size = 16),  
    plot.margin = margin(10, 10, 10, 10)
  )

var_fig_together + ggtitle("Simple model")
goby_out_fig_together + ggtitle("Original model")
```
